Whittaker Models for Real Groups

### 2024-06-24 whittaker.tex
updated last (SL2) section

Suppose $\xi_\R\in \Op\cap i\g(\R)^\ast$, and write $\xi_\R = F_\R^\ast$ where $F_\R$ is a principal nilpotent element. We may choose $H_\R, E_\R \in \g$ such that $(H_\R,E_\R,F_\R)$ is an $\mathrm{SL}(2)$-triple and  $E_\R$  contained in $i\g(\R)$,
which implies $H_\R\in \g(\R)$. 
After conjugating by~$G(\R)$ we may further assume $\theta(E_\R)=-F_\R$. Set 
\[   F_\theta=\frac12(-E_\R+F_\R+H_\R).\]
Then the Sekiguchi correspondence takes the $G(\R)$-orbit of~$\xi_\R$ to the $K$-orbit of $\xi_\theta=F_\theta^\ast$.

Computing the map in the other direction goes as follows.
Suppose $\xi_\theta \in \Op\cap \s^\ast$, and let $F_\theta \in \s$ be such that $\xi_\theta=F_\theta^\ast$. Choose $E_\theta, H_\theta \in \g$ such that $(E_\theta, H_\theta, F_\theta)$ is an $SL(2)$-triple with $E_\theta \in \s$,  which implies $H_\theta\in\k$. After conjugating by $K$ we may further assume $\sigma(E_\theta)=F_\theta$.
Define 
\begin{equation}\label{def_F_R}
  F_\R=\frac12(-E_\theta+F_\theta-H_\theta)\\
\end{equation}
Then the Sekiguchi correspondence takes the $K$-orbit of $\xi_\theta$ to the $G(\R)$-orbit of $\xi_\R = F_\R^\ast$.

@@

We know that  $\AV(\pi)=K \cdot F_\pi$, so the image of $\AV(\pi)$ by the Sekiguchi correspondence is 
 $G(\R)\cdot F_\R$. We want to check that this is equal to $\WF(\pi)$.

@@

Then $\Wh(\pi)=\Wh(\pi')$ if and only if $\mathscr{C}, \mathscr{C}'$ are in the same $W_K$-orbit. 
\end{lemma}

@@

>>>>

Suppose $\xi_\R\in \Op\cap i\g(\R)^\ast$, and write $\xi_\R = E_\R^\ast$ where $E_\R$ is a principal nilpotent element. We may choose $H_\R, E_\R \in \g$ such that $(H_\R,E_\R,F_\R)$ is an $\mathrm{SL}(2)$-triple and  $E_\R$  contained in $i\g(\R)$,
which implies $H_\R\in \g(\R)$. 
After conjugating by~$G(\R)$ we may further assume $\theta(E_\R)=-F_\R$. Set 
\[   E_\theta=\frac12(E_\R-F_\R+H_\R)\quad
F_\theta=\frac12(-E_\R+F_\R+H_\R)
\]
Then $E^\ast_\theta,F^\ast_\theta\in \Op\cap \s$, and the Sekiguchi correspondence takes the $G(\R)$-orbit of~$E^*_\R$ to the $K$-orbit of $E_\theta^\ast$. 
Equivalently, it takes the $G(\R)$-orbit of $F^*_\R$ to the $K$-orbits of $F_\theta^\ast$.

Computing the map in the other direction goes as follows.
Suppose $\xi_\theta \in \Op\cap \s^\ast$, and write $\xi_\theta=E_\theta^\ast$. Choose $F_\theta, H_\theta \in \g$ such that $(E_\theta, H_\theta, F_\theta)$ is an $SL(2)$-triple with $F_\theta \in \s$,  which implies $H_\theta\in\k$. After conjugating by $K$ we may further assume $\sigma(E_\theta)=F_\theta$.
Define 
\begin{equation}\label{def_F_R}
E_\R=\frac12(E_\theta-F_\theta-H_\theta)\quad
F_\R=\frac12(-E_\theta+F_\theta-H_\theta).
\end{equation}
Then the Sekiguchi correspondence takes the $K$-orbit of $E^*_\theta$ (resp. $F^*_\theta$) to the $G(\R)$-orbit of $E_\R^\ast$ (resp. $F_\R^\ast$).

@@

We know that  $\AV(\pi)=K \cdot F^*_\pi$, so the image of $\AV(\pi)$ by the Sekiguchi correspondence is 
 $G(\R)\cdot F_\R$. We want to check that this is equal to $\WF(\pi)$.

@@

Then $\Wh(\pi)=\Wh(\pi')$ if and only if $\mathscr{C}, \mathscr{C}'$ are conjugate by $W_K$.
\end{lemma}

@@

\subsection{The example of $\text{SL}_2$}

We illustrate the constructions of the previous section the case of $G=\SL(2,\R)$.

The complexified Lie algebra $\g$ is the space of $2x2$ complex matrices of trace $0$. Let $\s(\C)=\{t_z\mid z\in\C\}\subset \g$ where
$$
t_z=\begin{pmatrix}0&z\\-z&0
\end{pmatrix}
$$
Then $\s(\R)=\{t_x\mid x\in\R\}$.

For $z\in \C$ define $\lambda_z\in \s(\C)^*$ by $\lambda_z(t_x)=xz$.
The positive root can be taken to be   $\alpha(t_z)=2iz$, i.e.
$\alpha=\lambda_{2i}$ so $\rho=\frac12\alpha=\lambda_i$.
Note that $\alpha(\ch\alpha)=2$ gives
$$
\ch\alpha=t_{-i}
$$
i.e.
$$
\langle \lambda_{ik},\ch\alpha\rangle=\lambda_{ik}(t_{-i})=k
$$
Viewing $\ch\alpha$ as a map from $\C$ to $\ch{\mathfrak t}$ this says
$$
\ch\alpha(z)=t_{-iz}\quad (z\in \C).
$$
Also $\ch\rho=t_{-i/2}$
and
$$
\ch\rho(-i)=t_{-\frac 12}=\begin{pmatrix}0&-\frac 12\\\frac 12&0
\end{pmatrix}
$$

The Killing form satisfies
$$
\kappa(t_x,t_y)=-8xy.
$$
Define $H_{\lambda}\in \s(\C)$ so that $\kappa(H_\lambda,t_x)=\lambda(t_x)$. Then $H_\lambda=t_w$ for some $w$; so
$$
\kappa(H_{\lambda_z},t_x)=\lambda_z(t_x)=zx,\quad \kappa(t_w,t_x)=-8xw
$$
gives
$$
H_{\lambda_z}=t_{-\frac z8}.
$$
In particular
$$
H_{\pm \rho}=H_{\lambda_{\mp i}}=\begin{pmatrix}0&-\frac {\mp i}8\\\frac {\mp i}8&0
\end{pmatrix}
$$
Let $\pi(\lambda_{ik})$ be the discrete series representation with Harish-Chandra parameter $\lambda_{ik}$ $(k\in \Z_{\ne 0})$, so
$$
H_{\pi(\lambda_{ik})}=H_{\lambda_{ik}}=
\begin{pmatrix}0&-\frac{ik}8\\\frac{ik}8&0
\end{pmatrix}
$$
For an $SL(2)$-triple we can take $\{X_\alpha,X_{-\alpha},t_{-i}\}$ where
$$
X_\alpha=\frac12\begin{pmatrix}1&i\\i&-1
\end{pmatrix}, \quad X_{-\alpha}=\overline{X_\alpha}, \quad t_{-i}=\begin{pmatrix}0&-i\\i&0
\end{pmatrix}
$$
That is
$$
[t_z,X_\alpha]=\alpha(t_z)X_\alpha=\lambda_{2i}(t_z)X_\alpha=2izX_\alpha
$$
In particular
$$
[t_{-i},X_\alpha]=2X_\alpha
$$

Let's compute $\AV(\pi(\lambda_{ik})$ using  \eqref{e:Fpi}.
Suppose $k>0$. Then $\langle \lambda_{ik},\ch\alpha\rangle=k$, so $\Delta^+=\{\alpha\}$, and
$F_{\pi_{ik}}=X_{-\alpha}$. 
By \eqref{e:Fpi} $\AV(\pi_{ik})$ is the $K$-orbit of $X_{-\alpha}^\ast$.
Thus in general
$$
\AV(\pi_{ik})=K\cdot X^*_{-\sgn(k)\alpha}
$$
Note: this only uses \eqref{e:Fpi}, which comes directly from \cite{AV1}.

Conjugating $H_{\pi(\lambda_{ik})}$ by $\mathrm{diag}(x,\frac 1x)$ takes it to
$\begin{pmatrix}0&-x^2\frac{ik}8\\\frac{ik}{x^2}&0
\end{pmatrix}$,
and taking the limit we see
$$
\WF(\pi(\lambda_{ik}))=\AC(G(\R)\cdot H_{\pi(\lambda_{ik}}))=
G(\R)\cdot \begin{pmatrix}0&-\sgn(k)i\\0&0
\end{pmatrix}
$$
Note: this only uses  \cite{HarrisHeOlafsson}.

Note that the $G(\R)$ orbit of $\begin{pmatrix}0&i\\0&0
\end{pmatrix}$ contains $\begin{pmatrix}0&i\R^+\\0&0
\end{pmatrix}$,
so in particular $G(\R)\cdot \begin{pmatrix}0&ix\\0&0
\end{pmatrix}=G(\R)\cdot \begin{pmatrix}0&iy\\0&0
\end{pmatrix}$ if and only if $\sgn(x)=\sgn(y)$.
In particular
$$
\WF(\pi_{ik})=\WF(\pi_{i\ell})\Leftrightarrow \sgn(k)=\sgn(\ell)
$$

Finally let's compute $\Wh(\pi_{\lambda_{ik}})$ in two ways.
First of all by Lemma \ref{lem:matumoto}, which comes directly from \cite[Theorem A]{matumoto}
$$
\Wh(\pi_{\lambda_{ik}})=\WF(\pi_{\lambda_{ik}})=G(\R)\cdot \begin{pmatrix}0&-\sgn(k)i\\0&0
\end{pmatrix}
$$

Alternatively we apply Proposition \ref{JeffTasho_criterion}, which comes from 
\cite{adams_kaletha}. Note that

$$
\O_{\pi_{\lambda_{ik}}}=G(\R)\cdot H_{\lambda_{ik}}=G(\R)\cdot \begin{pmatrix}0&-\frac{ik}8\\\frac{ik}8&0
\end{pmatrix}
$$

Note that
$$
\K
\begin{pmatrix}0&\pm i\\0&0
\end{pmatrix}=\{\begin{pmatrix}0&\pm i\\z&0
\end{pmatrix}\mid z\in\C\}
$$
so, with $\epsilon=\pm 1$, 
$$
H_{\lambda_{ik}}\in \K\begin{pmatrix}0&\epsilon i\\0&0
\end{pmatrix}\Leftrightarrow \epsilon=\sgn(k).
$$

By Proposition \ref{JeffTasho_criterion}
we conclude (again) that
$$
\Wh(\pi_{\lambda_{ik}})=G(\R)\cdot\begin{pmatrix}0&-\sgn(k)i\\0&0
\end{pmatrix}
$$
